curatedMetagenomicData

### 2019-10-06 curatedMetagenomic_NS.Rmd
updated

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)

@@

?combined_metadata
View(combined_metadata)
table(combined_metadata$disease)

@@

table(sample_data(pseq.subset)$mnd)

table(sample_data(pseq.subset)$country)

```
## Check OTUs/Phylums
Prune(remove) taxa because we want to remove OTUs/taxa that aren't in any of the samples
Don't need to prune this taxa because OTUs are found in all samples.
Check this everytime you use the subset_samples function because of the pseq object being filtered. 
```{r check OTUs}
# Check if any OTUs are not present in any samples
any(taxa_sums(pseq.subset) == 0)
# FALSE, therefore, OTUs are found in all samples and we don't need to prune the taxa. If the answer is TRUE, some OTUs are not present in any samples. This is usually the case when data is subset to remove some samples. OTUs unique to those sample are not removed along with the samples. Therefore, it is important to check this everytime the phyloseq object is filtered for samples using `subset_samples` function.

@@

# Remove   everything without a country
pseq.subset <- subset_samples(pseq, !is.na(mnd))
pseq.subset <- subset_samples(pseq, disease == "healthy")

keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
f.pseq <- subset_taxa(pseq, keepotu)

@@

genus.pseq <- subset_taxa( pseq.subset, !is.na(Genus))
phylum.pseq <- subset_taxa( pseq, is.na(Class) & !is.na(Phylum))

taxonomy <- tax_table(genus.pseq)
# Pick the core (>0.1% relative abundance in >50% of the samples)
pseq.subsetc <- core(pseq.subset, detection = 0.1/100, prevalence = 50/100)

map_levels(NULL, "Phylum","Genus",pseq.c)

diversities(phylum.pseq, index = "all")

png("alpha.png")

tab <- diversities(pseq.c, index = "all")
stargazer(tab, type = "text",summary = NULL)

tab <- diversities(pseq.subsetc, index = "Shannon")
tab <- diversities(phylum.pseq, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
plot(tab)
dev.off()

# Get metadata
pseq.meta <- meta(phylum.pseq)
kable(head(pseq.meta))
stargazer(pseq.meta, type = "text",summary = NULL)

# Add the diversity table to metadata
pseq.meta$Shannon <- tab$shannon 
pseq.meta$InverseSimpson <- tab$inverse_simpson
pseq.meta$fisher <- tab$fisher
pseq.meta$chao1 <- tab$chao1
#wt = UniFrac(pseq.csubset, weighted=TRUE, normalized=FALSE, 
#             parallel=FALSE, fast=TRUE)
#plot(hclust(wt), main="Weighted UniFrac distances")

library(ggpubr)

p1 <- ggviolin(pseq.meta, x = "mnd", y = "Shannon",
 add = "boxplot", fill = "mnd", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
  print(p1)
ggsave("nation_violin_shannon.png")

keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
summary(keepotu)

fpseq <- subset_taxa(pseq, keepotu)

# This doesnt run. 
# plot_heatmap(fpseq, method="PCoA", distance="bray")

#keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
#summary(keepotu)

# subset_taxa(loman.pseq, keepotu)
#lfilt <- subset_taxa(pseq, keepotu & !is.na(Strain))
#plot_heatmap(lfilt, method="PCoA", distance="bray")

#lfilt <- subset_taxa(pseq, is.na(Class) & !is.na(Phylum))

#any(taxa_sums(lfilt) == 0)

# Alpha diversity again
png("alpha.png")

barplot(sort(taxa_sums(pseq.csubset), TRUE)[1:20] / nsamples(pseq.csubset),
        ylab = "Total counts", las = 2)

tab <- diversities(pseq.csubset)

tab <- diversities(pseq.subset, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
plot(tab)
dev.off()

# Get metadata
pseq.meta <- meta(pseq)
kable(head(pseq.meta))
stargazer(pseq.meta, type = "text",summary = NULL)

# Add the diversity table to metadata
pseq.meta$Shannon <- tab$shannon 
pseq.meta$InverseSimpson <- tab$inverse_simpson
pseq.meta$fisher <- tab$fisher
pseq.meta$chao1 <- tab$chao1

>>>>

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)
library(stargazer)
library(ggpubr)

@@

@@

table(sample_data(pseq)$mnd)

table(sample_data(pseq)$country)

```
## Check OTUs/Phylums
Prune(remove) taxa because we want to remove OTUs/taxa that aren't in any of the samples
Don't need to prune this taxa because OTUs are found in all samples.
Check this everytime you use the subset_samples function because of the pseq object being filtered. 
```{r check OTUs}
# Check if any OTUs are not present in any samples
any(taxa_sums(pseq) == 0)
# FALSE, therefore, OTUs are found in all samples and we don't need to prune the taxa. If the answer is TRUE, some OTUs are not present in any samples. This is usually the case when data is subset to remove some samples. OTUs unique to those sample are not removed along with the samples. Therefore, it is important to check this everytime the phyloseq object is filtered for samples using `subset_samples` function.

@@

# Remove   everything without a country

keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
pseq.f <- subset_taxa(pseq, keepotu)

pseq.c <- core(pseq, detection = 0.1/100, prevalence = 50/100)

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn")

# These are useful is filtering. 
#sample_variables(pseq.subset)
#get_variable(pseq.subset, sample_variables(pseq.subset)[4])

@@

# run the diversities. 	keep an eye on this as only one is working
tab <- diversities(pseq.subset, index = "Shannon")

# Get metadata
pseq.meta <- meta(pseq.subset)
kable(head(pseq.meta))
stargazer(pseq.meta, type = "text",summary = NULL)

# Add the diversity table to metadata
pseq.meta$Shannon <- tab$shannon 
#pseq.meta$InverseSimpson <- tab$inverse_simpson
#pseq.meta$fisher <- tab$fisher
#pseq.meta$chao1 <- tab$chao1

p1 <- ggviolin(pseq.meta, x = "mnd", y = "Shannon",
 add = "boxplot", fill = "mnd", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
  print(p1)
#ggsave("nation_violin_shannon.png")

# THis subsets the  genus and phylum. 
#genus.pseq <- subset_taxa( pseq.subset, !is.na(Genus))
#phylum.pseq <- subset_taxa( pseq, is.na(Class) & !is.na(Phylum))
